### 2023-02-02 markdown/Thesis.Rmd
structure of repo changed

---
title: " "
output: 
  pdf_document:
    number_sections: TRUE
geometry: "left = 2.5cm, right = 2cm, top = 2cm, bottom = 2cm"
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \usepackage{sectsty}
  - \usepackage{paralist}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \usepackage{dcolumn}
  - \usepackage{natbib}\bibliographystyle{agsm}
  - \usepackage[nottoc, numbib]{tocbibind}
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
#designed with help of: 
#https://ourcodingclub.github.io/tutorials/rmarkdown-dissertation/

knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)

#install dependencies
if (!requireNamespace("tinytex", quietly=TRUE))
  install.packages("tinytex")

library(tinytex)
```

\allsectionsfont{\centering}
\subsectionfont{\raggedright}
\subsubsectionfont{\raggedright}

\pagenumbering{gobble}

\begin{centering}

\vspace{3cm}
```{r uni_logo, echo=F, out.width="20%"}
knitr::include_graphics("FHCW_logo.png")
knitr::include_graphics("KI_logo.png")
```

\vspace{1cm}

\Large
\doublespacing
{\bf Design and implementation of analysis pipeline for single cell type proteomics data}

\vspace{1 cm}

\normalsize
\singlespacing
By

\vspace{0.5 cm}

\Large

{\bf Lukas Gamp}

\vspace{1.5 cm}

in partial fulfilment of the requirement \\for the degree of MSc with Honours \\in Bioinformatics

\vspace{1.5 cm}

\normalsize
mm yy
\end{centering}

\newpage

\pagenumbering{gobble}

\begin{centering}

{\bf Abstract}

\end{centering}

\spacing{1.5}

(the spacing is set to 1.5)

no more than 250 words for the abstract

- a description of the research question/knowledge gap – what we know and what we don’t know
- how your research has attempted to fill this gap
- a brief description of the methods
- brief results
- key conclusions that put the research into a larger context

\pagenumbering{roman}

\newpage

\centering
\raggedright
\newpage
\tableofcontents

\newpage

```{r acknowledgments, child='acknowledgments.Rmd'}
```

\newpage

```{r intro, child = 'introduction.Rmd'}
```
\pagenumbering{arabic}

\newpage

```{r methods, child = 'methods.Rmd'}
```

\newpage

```{r results, child = 'results.Rmd'}
```

\newpage

```{r discussion, child = 'discussion.Rmd'}
```

\newpage

```{r conclusion, child = 'conclusion.Rmd'}
```

\newpage

```{r biblio, child = 'bibliography.Rmd'}
```

\newpage

```{r appendix, child = 'appendix.Rmd'}
```