### 2023-02-02 markdown/acknowledgments.Rmd
child files added

---
title: ''
output: pdf_document
---

\section*{Acknowledgements}

Thank you for following this tutorial!

I hope you'll find it useful to write a very professional dissertation.