### 2023-02-02 markdown/appendix.Rmd

---
title: ''
output: pdf_document
---
# Appendix(ces)

## Appendix A: additional tables

Insert content for additional tables here.

\newpage

## Appendix B: additional figures

Insert content for additional figures here.

\newpage

## Appendix C: code

Insert code (if any) used during your dissertation work here.