### 2023-06-14 markdown/results.Rmd
results started

---
title: ''
output: pdf_document
---

# Results

After runring MaxQuant mode with MS2 search for TMTpro reporter ions with a runtime of 6 hours, the data was analyzed using the developed shiny application. 
Using the application starts with uploading the tabdelimited evidence and sample annotation file. The application is ready to run with default values. Fine tuning will be done subsequently.

```{r ui_summary, echo=F, out.width="100%", fig.align='center', fig.cap="Summary showing the process of the analysis pipeline"}
knitr::include_graphics("screenshots/summary.png")
```

On the left side of the user interface the analysis pipeline settings ca be changed. With the update button the user can apply the changes made and observe alterations in the dataset directly. The summary visualization looks cluttered since 36 files were processed, in smaller datasets the nodes of the network graph can be distuingished clearly.

By running the application with SCoPE2 \citep{Specht2021} reference cutoffs the following results were visualized.
```{r ui_numbers_default, echo=F, out.width="100%", fig.align='center', fig.cap="Barchart showing the number of peptide spectrum matches, peptides and proteins"}
knitr::include_graphics("screenshots/numbers_default.png")
```

The default settings yield a total number of unique proteins below $\approx 10^{3}$ aggregated from $10^{4.5} \approx$ peptide spectrum matches also aggregated over $10^{3.5}\approx$ peptides. The aggregation is set fixed by median.

```{r ui_mutual_info_default, echo=F, out.width="100%", fig.align='center', fig.cap="Mutual information with default setting"}
knitr::include_graphics("screenshots/mutual_info_default.png")
```

The mutual information (MI) within the sample type of the healthy control group shows a reduction to the lower edge of the IQR after the transformation and normalization. When comparing the MI with the sample type HIV with metabolic syndrome, the reduction is within the inter quantile range (IQR = 25 - 75 quantile) of the raw data MI.

```{r ui_RI_channel_default, echo=F, out.width="100%", fig.align='center', fig.cap="Reporter ion intensity with selection on channel. The selected factor will be highlighted and indicated in the legend."}
knitr::include_graphics("screenshots/RI_channel_default.png")
```

When comparing the median reporter ion intensities over channels show no systematic bias. Considering the 3 sample types occupy in every run the same channels, it is important to look at this visualization. Since the IQR overlap over the majority of samples, the change in expression with the factor channel can be considered as non-significant. However the channel will be later on tested for significance in the dimensionality redution plot and statistic module.

```{r ui_RI_file_default, echo=F, out.width="100%", fig.align='center', fig.cap="Reporter ion intensity with selection on file"}
knitr::include_graphics("screenshots/RI_file_default.png")
```

Comparing median reporter ion intensities with the focus on the batch (also Raw.file), it is obvious that the changes in signal are significant and need to be considered in further analysis. For the dimensionality reduction the user can choose in the sidepane the batch correction option ComBat \citep{Leek2012}. However this option is not chosen as default, since the conservative method includes the batch factor in the statistic module.

```{r ui_RI_st_default, echo=F, out.width="100%", fig.align='center', fig.cap="Reporter ion intensity with selection on sampletype"}
knitr::include_graphics("screenshots/RI_st_default.png")
```

Observing the Median RIs with a focus on sample types, IQRs partly overlap so the binding efficiency of the TMTpro reagent is not affected by the sample type.

```{r ui_cv_st_default, echo=F, out.width="100%", fig.align='center', fig.cap="Reporter ion intensity with selection on sampletype"}
knitr::include_graphics("screenshots/cv_st_default.png")
```

```{r ui_mutual_info_finetune, echo=F, out.width="100%", fig.align='center', fig.cap="Mutual information with adjusted settings"}
knitr::include_graphics("screenshots/mutual_info_finetune.png")
```