### 2023-06-21 markdown/acknowledgments.Rmd
proofread

---
title: ''
output: pdf_document
---

\section*{Acknowledgements}

In addition to my faithful feline companion, Apollo, who took care of my place in Austria while I was working on this project in Stockholm, I am deeply grateful to all the members of Ujjwal Neogi's research group.

Their unwavering guidance and support throughout this project have been truly invaluable, and I am profoundly thankful for their expertise and assistance.