### 2023-07-18 markdown/appendix.Rmd
headlines added. appendix moved to methods

---
title: ''
output: pdf_document
---
# Appendix

## Dry lab
### R-version info

```{r load_libraries, message=FALSE, warning=FALSE, include=FALSE}
library("shiny")
library("shinyWidgets")
library("shinydashboard")
library("DT")
library("spsComps")

library("scp")
library("SingleCellExperiment")
library("ggplot2")
library("magrittr")
library("dplyr")
library("reshape2")
library("scater")
library("limma")
library("CONSTANd")
library("stats")
library("impute")
library("sva")
library("plotly")
library("clusterProfiler")
library("infotheo")
library("org.Hs.eg.db")
library("AnnotationDbi")
library("piano")
library("snowfall")
library("parallel")
library("visNetwork")
```

```{r session_info, echo=FALSE}
sessionInfo()
```