### 2023-07-27 markdown/article.Rmd
article started

---
title: "article"
author: "Lukas Gamp"
date: "`r Sys.Date()`"
output: pdf_document
---
---
title: " "
output:
  bookdown::pdf_document2:
    toc: false
#    number_sections: TRUE
geometry: "left = 2.5cm, right = 2cm, top = 2cm, bottom = 2cm"
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \let\origfigure\figure
  - \let\endorigfigure\endfigure
  - \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
    } {
    \endorigfigure
    }
  - \usepackage{sectsty}
  - \usepackage{paralist}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \usepackage{dcolumn}
  - \usepackage{natbib}\bibliographystyle{agsm}
  - \usepackage[nottoc, numbib]{tocbibind}

bibliography: bibliography.bib
---

```{r setup, include=FALSE}
#designed with help of: 
#https://ourcodingclub.github.io/tutorials/rmarkdown-dissertation/

knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)

#install dependencies
if (!requireNamespace("tinytex", quietly=TRUE))
  install.packages("tinytex")

library(tinytex)

#install dependencies
if (!requireNamespace("DiagrammeR", quietly=TRUE))
  install.packages("DiagrammeR")

library(DiagrammeR)
```