### 2023-09-11 markdown/acknowledgments.Rmd

---
title: ''
output: pdf_document
---

\section*{Acknowledgements}

Beside my faithful feline companion, Apollo, who took care of my place in Austria while I worked on this project in Stockholm, I extend my deep gratitude to all the members of Ujjwal Neogi's research group. Their consistent guidance and support throughout this project have been truly invaluable, and I am profoundly appreciative of their expertise and assistance.

Furthermore, I wish to express my gratitude to Gerhard Dürnberger, my internal supervisor, for his unwavering support and feedback on my master thesis, along with his precious time devoted to our meetings.